---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---

```{r}
library(rnoaa)
library(tidyverse)
library(itsmr)
library(fpp2)
library(forecast)
library(urca)
```

```{r}
stataions<-ghcnd_stations()
```

```{r}
get_temperature <- function (stationid) {
  ghcnd_search(stationid = stationid, var = c("tmax"), 
               date_min = "1920-01-01", date_max = "2023-02-20")[[1]] %>%
  mutate(year = as.integer(format(date, "%Y")),
         month = as.integer(strftime(date, '%m')) %% 12, # make December "0"
         season = cut(month, breaks = c(0, 2, 5, 8, 11),
                      include.lowest = TRUE,
                      labels = c("Winter", "Spring", "Summer", "Fall")),
         year = if_else(month == 0, year + 1L, year)) %>%
  group_by(year, season) %>%
  summarize(tmax_avg = mean(tmax, na.rm = TRUE))
}
```

```{r}
liestal1<-get_temperature("GME00127786")
```

```{r}
winterw<-liestal1%>%filter(season=="Spring")
head(winterw)

winterw
```



```{r}
liestal_raw<- read.csv("data/liestal.csv")

liestal_1950<-liestal_raw[liestal_raw$year>=1950 & liestal_raw$year<=2022,]
head(liestal_1950)
```

```{r}
newdata<-left_join(winterw,liestal_1950,by="year")

newdata
```

# mod1.1 spring max tempreture vs bloom_day

```{r}
mod1.1=lm(bloom_doy~tmax_avg,data=newdata)
summary(mod1.1)
```


# mod1.2 arima ts model for residules form mod1.1
```{r}
liestal.ts22<-ts(data=liestal_raw$bloom_doy[liestal_raw$year>=1950],start=1954) # raw data for doy
plot.ts(liestal.ts22,main="bloom_doy")
residts=ts(mod1.1$residuals,start=1954)
plot.ts(residts,main="residual_ts for mod1.1")

acf(residts,main="acf for residual from mod1.1")
```


```{r}
mod1.2=auto.arima(residts)
summary(mod1.2)
```

```{r}
forecast(mod1.2,5)
```

```{r}
for1=forecast(mod1.2,5)
plot(forecast(mod1.2,5))

for1
```

# mod 1.3 arima ts model for max temperature

use the point forcasted 
```{r}
head(newdata)

l.ts23<-ts(newdata$tmax_avg[newdata$year>=1950],start = 1950)
plot(l.ts23,main="temp")
acf(l.ts23) 
```

# holt's method has smaller training error, use holt's method 

```{r}
fc <- holt(l.ts23, damped=TRUE,h=10) # use holt's method 

fc3<-auto.arima(l.ts23) # use autoarima
```

```{r}
fc
autoplot(fc)
forecast(fc3,h=10)
plot(forecast(fc3,h=10))
accuracy(fc)
accuracy(fc3)
tempfc2=fc[[2]]
tempfc2
```

# final prediction choose this version, since temp captures someamount of variability, the residual seems not depend on time, holt's linear method for predicting the tempture may be resonable.

```{r}
tempfc2_m=as.matrix(tempfc2)


tempfc2_m
```

```{r}

mod1.1forecast <- mod1.1$coefficients[1]+mod1.1$coefficients[2]*tempfc2_m
mod1.1forecast
```

```{r}
forecast(mod1.2,10)
mod1.2forcast=forecast(mod1.2,10)[[4]][1:10]
```

final prediction for KYOTO 1 
```{r}


lfinalpred=round(mod1.2forcast+mod1.1forecast)

lfinalpred
```

```{r}
comb=mod1.1$fitted.values+mod1.2$fitted
plot.ts(liestal.ts22,xlim=c(1950, max(time(liestal.ts22))+10))
tyears = ((max(time(liestal.ts22))+1):(max(time(liestal.ts22))+10))
lines(tyears,lfinalpred,col="red") 
lines(comb,col="red")
```


# second try:
```{r}
l_1950<-liestal_raw[liestal_raw$year>=1950 & liestal_raw$year<=2022,"bloom_doy"]
l_1950_ts<-ts(data=l_1950,start=1950)
```

```{r}
acf(l_1950_ts,20, ylim=range(-1,1))
```

```{r}
mod2<-auto.arima(l_1950_ts)
```


```{r}
forecast(mod2,10)

plot(forecast(mod2,10))
```

```{r}
train <- window(l_1950_ts, end=c(2010))
h <- length(l_1950_ts) - length(train)
ETS <- forecast(ets(train), h=h)
ARIMA <- forecast(auto.arima(train),
  h=h)
HOLT<-holt(train,damped=TRUE,h=h)
NNAR <- forecast(nnetar(train), h=h)
TBATS <- forecast(tbats(train, biasadj=TRUE), h=h)


autoplot(l_1950_ts) +
  autolayer(ETS, series="ETS", PI=FALSE) +
  autolayer(ARIMA, series="ARIMA", PI=FALSE) +
  autolayer(HOLT,series="HOLT",PI=FALSE)+
  autolayer(NNAR, series="NNAR", PI=FALSE) +
  autolayer(TBATS, series="TBATS", PI=FALSE) +
  
 
  ggtitle("resids")
```


